---
title: "Brisbane polling booths"
description: |
  Analysis of Brisbane polling booths and thier progressiveness.
author:
  - name: Paul Melloy 
    url: https://github.com/PaulMelloy/fusion2022
    affiliation: Fusion party 
    affiliation_url: https://www.fusionparty.org.au/
date: "`r Sys.Date()`"
output: distill::distill_article
---

## Load libraries
Loading 
```{r librarys}
library(eechidna)
library(tidyverse)
library(ggplot2)
library(ggalluvial)
```

## Inspect 2019 two party prefered  

```{r}
tpp19 %>%
  filter(StateAb == "QLD") %>%
  head()
```

## Inspect 2019 first preferences  

```{r}
fp19 %>%
  filter(StateAb == "QLD",
         DivisionNm == "RYAN") %>%
  arrange(-Percent)
```


## Polling booth preferences  

Peek at the data  

```{r download_firstPrefBooth}
fppb <- firstpref_pollingbooth_download()
head(fppb)
```

Which polling booth in the Ryan division had the highest votter turnout in 2016

```{r}
fppb %>%
  filter(DivisionNm == "RYAN",
         year == 2016) %>%
  group_by(PollingPlace) %>%
  summarise(voter_turnout = sum(OrdinaryVotes, na.rm = TRUE)) %>%
  arrange(-voter_turnout)
```
This R package does not have the 2019 election data so we will download it for QLD from the [AE website](https://results.aec.gov.au/24310/Website/HouseDownloadsMenu-24310-Csv.htm.)  

```{r}
fp19_qld <- read.csv("https://results.aec.gov.au/24310/Website/Downloads/HouseStateFirstPrefsByPollingPlaceDownload-24310-QLD.csv",skip = 1)

head(fp19_qld)
```

```{r}
fp19_qld %>%
  filter(DivisionNm == "Ryan") %>%
  group_by(PollingPlace) %>%
  summarise(voter_turnout = sum(OrdinaryVotes)) %>%
  arrange(-voter_turnout)
```


Add a progressiveness score to each party

```{r}
fp19_qld <-
  fp19_qld %>%
  mutate(progresive =
           case_when(
             PartyAb == "FACN" ~ 0,
             PartyAb == "ALP" ~ 65,
             PartyAb == "ON" ~ 20,
             PartyAb == "LNP" ~ 45,
             PartyAb == "UAPP" ~ 25,
             PartyAb == "DLP" ~ 50,
             PartyAb == "GRN" ~ 80,
             PartyAb == "SAL" ~ 85,
             PartyAb == "KAP" ~ 50,
             PartyAb == "AJP" ~ 75,
             PartyAb == "LDP" ~ 23,
             PartyAb == "SPP" ~ 70,
             PartyAb == "LAOL" ~ 15,
             PartyAb == "AUP" ~ 77,
             PartyAb == "AFN" ~ 35,
             PartyAb == "SEP" ~ 70,
             TRUE ~ NA_real_))
```

Which booth had the higher average progressive vote

```{r}
fp19_qld %>%
  filter(DivisionNm == "Ryan") %>%
  group_by(PollingPlace) %>%
  summarise(progresiveness = sum(progresive * OrdinaryVotes, na.rm = TRUE),
            votes = sum(OrdinaryVotes, na.rm = TRUE)) %>%
  mutate(progresivenessP = progresiveness / votes) %>%
  arrange(-progresivenessP) %>%
  print(n = 46)
```



Lets plot the flow of preferences on each counting round.

```{r}

dop_qld_zip <- paste0(tempfile(),".zip")
download.file(url = "https://results.aec.gov.au/24310/Website/External/HouseDopByPPDownload-24310-QLD.zip",
              destfile = dop_qld_zip)

dir.create("data/dop_qld")

unzip(zipfile =  dop_qld_zip, exdir = "data/dop_qld")

dop_ryan <- read.csv("data/dop_qld/HouseDopByPPDownload-24310-QLD-RYAN.csv",
                      skip = 1)
# polling booths
unique(dop_ryan$PPNm)

# get custom script
source("R/plot_preference_flow.R")

plot_preference_flow(dop_ryan,
                     division = "Ryan",
                     polling_booth = "The Gap West")


```


```{r}
plot_preference_flow(dop_ryan,
                     division = "Ryan",
                     polling_booth = "Toowong")
```

```{r}
plot_preference_flow(dop_ryan,
                     division = "Ryan",
                     polling_booth = "Bellbowrie")
```